A user-extensible and adaptable parser architecture

### 2008-06-02 papers/sgai-2008/paper.tex
Corrections after final proofreading.

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental measures where there is
a deficiency.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms},
which requires accepting the mail in the first place, but rejecting is
preferable in many cases: senders of non-spam mail that is mistakenly
rejected will receive an immediate non-delivery notice; load is reduced on
mail servers (allowing more intensive content-based measures to be used);
users have less mail to wade through.  Improving the performance of
anti-spam measures which are applied when mail is being transferred via
SMTP\footnotemark{} is the goal of this project, by providing a platform
for reasoning about anti-spam filters.  The approach chosen to measure
performance was to analyse the log files produced by the mail server,
rather than modifying the mail server to generate performance statistics;
this approach improves the chances of other sites testing and utilising the
software.  The need arose for a parser capable of dealing with the great
number and variety of log lines produced by Postfix~\cite{postfix}, the
SMTP server in use.  The parser must be designed so that adding support for
parsing new log lines is a simple task, as the log lines will continue to
evolve over time.  The variety in log lines occurs for several
reasons:\footnotetext{Simple Mail Transfer Protocol transfers mail from
sender to one or more recipients across the Internet.  It is a simple,
human readable, plain text protocol, making it quite easy to test and debug
problems with it.  The original protocol definition is
RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

Decoupling the parsing rules from the associated actions and framework
allows new rules to be written and tested without requiring modifications
to the parser source code, significantly lowering the barrier to entry for
casual users who need to parse new log lines, e.g.\ part-time systems
administrators attempting to reduce spam.  Decoupling the actions from the
framework simplifies both framework and actions: the framework provides
services to the actions, and doesn't need to perform any tasks specific to
the input being parsed; actions benefit from having services provided by
the framework, freeing them to concentrate on the task of accurately and
correctly processing the information provided by rules.

Decoupling also creates a clear separation of functionality: rules handle
low level details of identifying and extracting data from log lines;
actions handle the higher level details of assembling the required data,
dealing with the intricacies of the input being parsed, complications
arising, etc; the framework provides services to actions and manages the
parsing process.

@@

The framework provides utility and support functions for the actions and
rules.  It provides shared storage to pass data between actions, most
notably the data structures used to represent each connection and mail.  It
loads and validates rules, controls parsing of log files, invokes actions,
tracks how often each rule matches (to optimise rule ordering, see
\sectionref{Rule ordering}), and stores results in the database.  The
framework is independent of the parser being written: the same
functionality will be required for most or all parsers.

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser will need to write the actions it requires from scratch, unless
it's extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  Unsurprisingly the action with the
highest number of associated rules is \texttt{REJECTION}, the action which
handles rejections; it's followed by \texttt{SAVE\_BY\_QUEUEID}, the action
responsible for handling informative inputs: these are log lines that
supplement the data gathered from other log lines (such as rejections), but
that don't by themselves represent interesting events.  The third most
common action is, perhaps surprisingly, \texttt{IGNORE}: there are many log
lines which are not of interest to the parser, e.g.\ warning messages about
DNS mismatches.  The parser warns about every unparsed log line, to alert
the user that they need to alter or extend their rules; the \texttt{IGNORE}
action does nothing when executed, allowing uninteresting log lines to be
parsed without causing any effects.  The remaining actions have only one or
two associated rules: in some cases this is because there will only ever be
one type of log line for that action, e.g.\ all log lines showing that a
remote client has connected are matched by a single rule and handled by the
\texttt{CONNECT} action; in other cases the action is required to address a
deficiency in the log files, or a complication which arises during parsing.

@@

The ability to add special purpose actions to deal with difficulties and
new requirements which are discovered during parser development is one of
the strengths of this architecture: instead of writing one monolithic
function which must be modified to support new behaviour or resolve
challenges, with all the attendant risks of adversely affecting the
existing, working parser, when a new requirement arises an independent
action can be written to satisfy it.  In some cases the new action will
require the cooperation of other actions, e.g.\ to set or check a flag;
there is a possibility of introducing failure by modifying existing
actions, but modifications will be smaller and occur less frequently than
with a monolithic architecture, and thus failures should be less likely and
easier to test for and diagnose.  The parser can be written in an object
oriented style, allowing sub-classes to extend or override actions in
addition to adding new actions; because each action is an independent
function, the sub-class need only modify the action it is interested in,
rather than reproducing large chunks of functionality.

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.  Some were the result of deficiencies in Postfix's logging,
(some of those deficiencies were improved or resolved by later versions of
Postfix); others were due to the vagaries of process scheduling, client
behaviour, and administrative actions.  All were successfully accommodated
in the parser: adding new actions was enough for several of the
complications; others required modifications to a single existing action to
work around the difficulties; the remainder were resolved by adapting
actions to pass information to subsequent actions (via the framework),
those subsequent actions then check for the extra information and change
their behaviour appropriately.

@@

Rules are responsible for categorising inputs, specifying both the data to
be extracted from each log line and the action that should be invoked.  The
Postfix log parser stores the rules in the same SQL database that the
results are stored in; this removes any doubt as to which set of rules was
used to produce a set of results.  Other implementations are free to store
the rules in whatever fashion suits their needs.  Each rule defines several
attributes:

@@

    \item [hits] This counter is used to optimise the order in which rules
        are tried against the input lines.  Assuming that the distribution
        of log lines is reasonably consistent between log files, rules
        matching more frequently will be tried before rules matching less
        frequently, lowering the parser's execution time.

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the first matching rule
wins and the order in which rules are tried against each line may change
between log files.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating SMTP delivery to specific
sites from SMTP delivery to the rest of the world.  The algorithm provides
a facility for specifying the order of overlapping rules: the priority
field in each rule (defaults to zero).  Rules are sorted by priority,
highest first, and then rules with the same priority are sorted by the
number of successful matches when parsing the previous log file.  Negative
priorities may be useful for catchall rules.

@@

% Don't reformat this!
% XXX CONSIDER USING A TABLE ENVIRONMENT; IT DIDN'T LOOK GREAT BEFORE.
\begin{tabular}[]{ll}
\textbf{Attribute}        & \textbf{Value}                                   \\
program                   & \daemon{smtpd}                                   \\
regex                     & \verb!^connect from (__HOSTNAME__)\[(__IP__)\]$! \\
result\_cols              &                                                  \\
connection\_cols~\empty{} & client\_hostname = 1; client\_ip = 2             \\
action                    & CONNECT                                          \\
priority                  & 0                                                \\
\end{tabular}

\noindent{}The \texttt{CONNECT} action is described in \sectionref{Actions}.

\subsection{Architecture characteristics}

@@

    \item [Partially context-sensitive actions and cascaded parsing:]
        Actions can consult the results (or lack of results) of previous
        actions when executing, providing some context sensitivity.
        Actions can not inspect past or upcoming inputs, however allowing
        actions can return a modified input line that will be parsed as if
        read from the input stream, allowing for a simplified version of
        cascaded parsing~\cite{cascaded-parsing}.

    \item [Matching rules to log lines is simple:]  The first matching rule
        wins: there is no backtracking to try alternate rules, no attempt
        is made to pick a \textit{best\/} rule; the first rule which
        matches the line is the rule which parses it and determines the
        action that will be invoked.  The priority attribute may be used to
        change the ordering of rules.

@@

    \item [Line oriented:]  The architecture is line oriented at present:
        there is no facility for rules to consume more input or push unused
        input back onto the input stream.  This was not a deliberate
        decision, rather a consequence of the line oriented nature of the
        log files; other implementations could pursue a more flexible
        approach.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 75 MB log files containing 300,000 log lines (generated daily on
a mail server handling mail for approximately 700 users --- large scale
mail servers would produce much larger log files on a daily basis).  When
generating the data for the graphs included in this document,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times, the results from first run discarded,
and the execution time for the remaining 9 runs averaged.

@@

\subsection{Rule ordering}

@@

The dips at log files 22 and 62--68 in fig.~\refwithpage{percentage
increase of shuffled and reversed over normal} correspond to peaks in log
file size in fig.~\refwithpage{Execution time, file size, and number of
lines}, and peaks in fig.~\refwithpage{execution time vs file size vs
number lines factor} (where a peak means that performance is better).  The
explanation for this took some time to arrive at, but it turns out to be
reasonably simple.  The large log files in question were caused by a mail
forwarding loop, where the distribution of log lines is quite different to
normal, resulting in different performance characteristics.  The vast
majority of log lines when a mail loop occurs are from Postfix components
which have a small number of rules associated with them, whereas in general
the majority of log lines are from the Postfix component (\daemon{smtpd})
with the highest number of rules.  Furthermore in normal log files the
majority of log lines are distributed much more evenly across rules than
the majority of log lines during a mail loop.

These two characteristics combine to reduce the average number or rules
required to parse a log line when there is a mail loop, as shown by the
peaks in fig.~\refwithpage{execution time vs file size vs number lines
factor} and troughs in fig.~\refwithpage{percentage increase of shuffled
and reversed over normal}.  When the rule ordering is reversed the majority
of log lines generated by a mail loop will be parsed using very few rules,
because there are very few rules for the Postfix components producing the
log lines.  Without a mail loop the Postfix components producing the
majority of log lines have a large number of rules associated with them,
and reverse ordering deliberately starts with the least likely rules and
moves towards the most likely.  This leads to a noticeable drop in the
average time required to parse a log line during a mail loop, as shown in
fig.~\refwithpage{percentage increase of shuffled and reversed over
normal}.  The number of rules which need to be consulted when the ordering
is shuffled varies between the optimal and the pessimal, with the
performance varying proportionally.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full ruleset, whereas the minimum number of
rules required to parse the \numberOFlogFILES{} log files is
\numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the total.  A
second set of statistics was generated using the minimum ruleset and the
results compared to the statistics generated using the full ruleset: the
percentage execution increase when using the full ruleset instead of the
minimal ruleset for normal, shuffled and reversed orderings is shown in
fig.~\refwithpage{Percentage execution time increase of maximum ruleset
over minimum ruleset}.  \showgraph{build/plot-normal-vs-smaller}{Percentage
execution time increase of maximum ruleset over minimum ruleset}{Percentage
execution time increase of maximum ruleset over minimum ruleset}

It is clear from fig.~\refwithpage{Percentage execution time increase of
maximum ruleset over minimum ruleset} that the increased number of rules
has a noticeable performance impact with reverse ordering, and to a lesser
extent with shuffled ordering.  The normal ordering (where are the most
frequently matching rules are tried first) shows a
\input{build/normal-vs-smaller-percentage-increase.tex} increase in
execution time for a \numberOFrulesMAXIMUMpercentage{} increase in the
number of rules.  These results show that the architecture scales very well
as the number of rules increases, and that sorting the rules is an
important optimisation enabling this.

@@

The architecture's greatest strength is the ease with which is can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new one with an appropriate regex and the task is complete; knowledge of
the framework is not required when writing rules.  Parsing a new category
of input is achieved by writing a new action and appropriate rules; quite
often the action will not need to interact directly with existing actions,
but when interaction is required the framework provides the necessary
facilities.  The decoupling of rules from actions allows different sets of
rules to be used with the same actions: e.g.\ a parser might have actions
to process versions one and two of a file format; the parser will then
parse either version one or version two depending on the ruleset selected.
The architecture makes it possible to apply commonly used programming
techniques such as object orientation, inheritance, composition, delegation
and roles when designing and implementing a parser, simplifying the process
of working within a team with shared responsibility or developing and
testing additional functionality.  This architecture is ideally suited to
parsing inputs where the input is not fully understood or does not follow a
fixed grammar: the architecture warns about unparsed inputs and other
errors, but continues on parsing as best it can, allowing the developer of
a new parser to decide which deficiencies are most important and require
attention first, rather than being forced to fix the first error which
arises.

The dataset gathered by the parser provides the foundation for the future
of this project: applying machine-learning algorithms to the data to
analyse and optimise the set of rules in use, followed by identifying
patterns in the data which could be used to write new filters to recognise
and reject spam rather than accepting it.  The current focus is on
clustering and decision trees to optimise the order in which rules are
applied; future efforts will involve using gathered data to train and test
new filters.  This task is very similar to optimising a black box
application based on its inputs and outputs, and this approach could be
applied to optimising the behaviour of any system given sufficient logging
to analyse.  An alternative approach to black box optimisation which uses
application profiling in conjunction with the application's error messages
to improve the error messages shown to users is described
in~\cite{black-box-error-reporting}; profiling data may be useful in
supplementing systems which fail to provide adequate log messages.

>>>>

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where there is
a deficiency.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms},
which requires accepting the mail in the first place, but rejecting is
preferable in many cases: senders of non-spam mail that is mistakenly
rejected will receive an immediate non-delivery notice; load is reduced on
mail servers (allowing more intensive content-based techniques to be used);
users have less mail to wade through.  Improving the performance of
anti-spam techniques which are applied when mail is being transferred via
SMTP\footnotemark{} is the goal of this project, by providing a platform
for reasoning about anti-spam filters.  The approach chosen to measure
performance was to analyse the log files produced by the mail server,
rather than modifying the mail server to generate performance statistics;
this approach improves the chances of other sites testing and utilising the
software.  The need arose for a parser capable of dealing with the great
number and variety of log lines produced by Postfix~\cite{postfix}, the
SMTP server in use.  The parser must be designed so that adding support for
parsing new log lines is a simple task, as the log lines will continue to
evolve over time.  The variety in log lines occurs for several
reasons:\footnotetext{Simple Mail Transfer Protocol transfers mail from the
sender to one or more recipients across the Internet.  It is a simple,
human readable, plain text protocol, making it quite easy to test and debug
problems with it.  The original protocol definition is
RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

Decoupling the parsing rules from their associated actions and the
framework allows new rules to be written and tested without requiring
modifications to the parser source code, significantly lowering the barrier
to entry for casual users who need to parse new log lines, e.g.\ part-time
systems administrators attempting to reduce spam.  Decoupling the actions
from the framework simplifies both framework and actions: the framework
provides services to the actions, but doesn't need to perform any tasks
specific to the input being parsed; actions benefit from having services
provided by the framework, freeing them to concentrate on the task of
accurately and correctly processing the information provided by rules.

Decoupling also creates a clear separation of functionality: rules handle
low level details of identifying and extracting data from log lines;
actions handle the higher level details of assembling the required data,
dealing with the intricacies of the input being parsed, complications
arising, etc.;\ the framework provides services to actions and manages the
parsing process.

@@

The framework provides utility and support functions for the actions and
rules.  It provides shared storage to pass data between actions, most
notably the data structures used to represent each connection and mail.  It
loads and validates rules, controls parsing of log files, invokes actions,
tracks how often each rule matches, to optimise rule ordering (see
\sectionref{Rule ordering}), and stores results in the database.  The
framework is independent of the parser being written: the same
functionality will be required for most or all parsers.

@@

Each action contains the code required to deal with log lines from a
particular category, e.g.\ rejections.  The actions are parser-specific:
each parser will need to write the actions it requires from scratch, unless
it's extending an existing parser.  In the Postfix parser developed for
this project there are \numberOFactions{} actions and \numberOFrules{}
rules; the distribution of rules to actions is very skewed, as shown in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  Unsurprisingly the action with the
highest number of associated rules is \texttt{REJECTION}, the action which
handles rejections; it's followed by \texttt{SAVE\_BY\_QUEUEID}, the action
responsible for handling informative inputs: these are log lines that
supplement the data gathered from other log lines (such as rejections), but
that don't by themselves represent interesting events.  The third most
common action is, perhaps surprisingly, \texttt{IGNORE}: there are many log
lines which are not of interest to the parser, e.g.\ warning messages about
DNS mismatches.  The parser warns about every unparsed log line, to alert
the user that they need to alter or extend their ruleset; the
\texttt{IGNORE} action does nothing when executed, allowing uninteresting
log lines to be parsed without causing any effects.  The remaining actions
have only one or two associated rules: in some cases this is because there
will only ever be one type of log line for that action, e.g.\ all log lines
showing that a remote client has connected are matched by a single rule and
handled by the \texttt{CONNECT} action; in other cases the action is
required to address a deficiency in the log files, or a complication which
arises during parsing.

\label{example action}

@@

The ability to add special purpose actions to deal with difficulties and
new requirements which are discovered during parser development is one of
the strengths of this architecture: instead of writing one monolithic
function which must be modified to support new behaviour or resolve
challenges, with all the attendant risks of adversely affecting the
existing parser, when a new requirement arises an independent action can be
written to satisfy it.  In some cases the new action will require the
cooperation of other actions, e.g.\ to set or check a flag; there is a
possibility of introducing failure by modifying existing actions, but
modifications will be smaller and occur less frequently than with a
monolithic architecture, thus failures should be less likely, and easier to
test for and diagnose.  The parser can be written in an object oriented
style, allowing sub-classes to extend or override actions in addition to
adding new actions; because each action is an independent function, the
sub-class need only modify the action it is interested in, rather than
reproducing large chunks of functionality.

@@

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.  Some were the result of deficiencies in Postfix's logging
(some of those deficiencies were improved or resolved by later versions of
Postfix); others were due to the vagaries of process scheduling, client
behaviour, and administrative actions.  All were successfully accommodated
in the parser: adding new actions was enough for several of the
complications; others required modifications to a single existing action to
work around the difficulties; the remainder were resolved by adapting
actions to pass information to subsequent actions (via the framework),
those subsequent actions then check for the extra information and change
their behaviour appropriately.

@@

Rules are responsible for categorising inputs, specifying both the data to
be extracted from each log line and the action that should be invoked.  The
Postfix log parser stores the rules in the same SQL database that the
results are stored in, removing any doubt as to which set of rules was used
to produce a set of results; other implementations are free to store the
rules in whatever fashion suits their needs.  Each rule defines several
attributes:

@@

    \item [hits] This counter is used to optimise the order in which rules
        are tried against the input lines.  Assuming that the distribution
        of log lines is reasonably consistent between log files, rules
        matching more frequently will be tried before rules matching less
        frequently, reducing the parser's execution time.

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the first matching rule
wins, and the order in which rules are tried against each line may change
between log files.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating SMTP delivery to specific
sites from SMTP delivery to the rest of the world.  The algorithm provides
a facility for specifying the order of overlapping rules: the priority
field in each rule (defaults to zero).  Rules are sorted by priority,
highest first, and then rules with the same priority are sorted by the
number of successful matches they had parsing the previous log file.  Negative
priorities may be useful for catchall rules.

@@

% Don't reformat this!
% XXX CONSIDER USING A TABLE ENVIRONMENT; IT DIDN'T LOOK GREAT BEFORE.
\begin{tabular}[]{ll}
    \hline
    \noalign{\smallskip}
    Attribute                 & Value                                            \\
    \noalign{\smallskip}
    \hline
    \noalign{\smallskip}
    program                   & \daemon{smtpd}                                   \\
    regex                     & \verb!^connect from (__HOSTNAME__)\[(__IP__)\]$! \\
    result\_cols              &                                                  \\
    connection\_cols~\empty{} & client\_hostname = 1; client\_ip = 2             \\
    action                    & CONNECT                                          \\
    priority                  & 0                                                \\
    \noalign{\smallskip}
    \hline
    \noalign{\smallskip}
\end{tabular}

\noindent{}The \texttt{CONNECT} action is described in \sectionref{example
action}.

\subsection{Architecture Characteristics}

@@

    \item [Partially context-sensitive actions and cascaded parsing:]
        Actions can consult the results (or lack of results) of previous
        actions when executing, providing some context sensitivity.
        Actions can not inspect past or upcoming inputs, however actions
        can return a modified input line that will be parsed as if read
        from the input stream, allowing for a simplified version of
        cascaded parsing~\cite{cascaded-parsing}.

    \item [Matching rules to log lines is simple:]  The first matching rule
        wins: there is no backtracking to try alternate rules, no attempt
        is made to pick a \textit{best\/} rule; the first rule which
        matches the line is the rule which parses it and determines the
        action that will be invoked.

@@

    \item [Line oriented:]  The architecture is line oriented at present:
        there is no facility for rules to consume more input or push unused
        input back onto the input stream.  This was not a deliberate
        decision, rather a consequence of the line oriented nature of the
        log files; more flexible approaches could be pursued.

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 75 MB log files containing 300,000 log lines (generated daily on
a mail server handling mail for approximately 700 users --- large scale
mail servers would produce much larger log files on a daily basis).  When
generating the data for the graphs included in this document,
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times, the results from the first run
discarded, and the execution time for the remaining 9 runs averaged.

@@

\subsection{Rule Ordering}

@@

The dips at log files 22 and 62--68 in fig.~\refwithpage{percentage
increase of shuffled and reversed over normal} correspond to peaks in log
file size in fig.~\refwithpage{Execution time, file size, and number of
lines}, and peaks in fig.~\refwithpage{execution time vs file size vs
number lines factor}, where a peak means that performance is better.  The
explanation for this took some time to arrive at, but it turns out to be
reasonably simple.  The large log files in question were caused by a mail
forwarding loop, where the distribution of log lines is quite different to
normal, resulting in different performance characteristics.  The vast
majority of log lines when a mail loop occurs are from Postfix components
which have a small number of rules associated with them, whereas in general
the majority of log lines are from the Postfix component (\daemon{smtpd})
with the highest number of rules.  Furthermore in normal log files the
majority of log lines are distributed much more evenly across rules than
the majority of log lines during a mail loop.

These two characteristics combine to reduce the average number of rules
required to parse a log line when there is a mail loop, as shown by the
peaks in fig.~\refwithpage{execution time vs file size vs number lines
factor} and troughs in fig.~\refwithpage{percentage increase of shuffled
and reversed over normal}.  When the rule ordering is reversed the majority
of log lines generated by a mail loop will be parsed using very few rules,
because there are very few rules for the Postfix components producing the
log lines.  Without a mail loop the Postfix components producing the
majority of log lines have a large number of rules associated with them,
and reverse ordering deliberately starts with the least likely rules and
moves towards the most likely.  This leads to a noticeable drop in the
average time required to parse a log line during a mail loop, as shown in
fig.~\refwithpage{percentage increase of shuffled and reversed over
normal}.  The number of rules which need to be consulted when the ordering
is shuffled varies between the optimal and the pessimal, with the
performance varying proportionally.

@@

How any architecture scales as the number of rules increases is important,
but it is particularly important in this architecture because it is
expected that there will be a large number of rules.  There are
\numberOFrules{} rules in the full Postfix ruleset, whereas the minimum
number of rules required to parse the \numberOFlogFILES{} log files is
\numberOFrulesMINIMUM{}, \numberOFrulesMINIMUMpercentage{} of the total.  A
second set of statistics was generated using the minimum ruleset and the
results compared to the statistics generated using the full ruleset: the
percentage execution increase when using the full ruleset instead of the
minimal ruleset for normal, shuffled and reversed orderings is shown in
fig.~\refwithpage{Percentage execution time increase of maximum ruleset
over minimum ruleset}.  \showgraph{build/plot-normal-vs-smaller}{Percentage
execution time increase of maximum ruleset over minimum ruleset}{Percentage
execution time increase of maximum ruleset over minimum ruleset}

It is clear from fig.~\refwithpage{Percentage execution time increase of
maximum ruleset over minimum ruleset} that the increased number of rules
has a noticeable performance impact with reverse ordering, and to a lesser
extent with shuffled ordering.  The normal ordering (where are the most
frequently matching rules are tried first) shows a mean increase of 
\input{build/normal-vs-smaller-percentage-increase.tex} in
execution time for a \numberOFrulesMAXIMUMpercentage{} increase in the
number of rules.  These results show that the architecture scales very well
as the number of rules increases, and that sorting the rules is an
important optimisation enabling this.

@@

The architecture's greatest strength is the ease with which it can be
adapted to deal with new requirements and inputs.  Parsing a variation of
an existing input is a trivial task: simply modify an existing rule or add
a new one with an appropriate regex and the task is complete; knowledge of
the framework is not required when writing rules.  Parsing a new category
of input is achieved by writing a new action and appropriate rules; quite
often the action will not need to interact directly with existing actions,
but when interaction is required the framework provides the necessary
facilities.  The decoupling of rules from actions allows different sets of
rules to be used with the same actions: e.g.\ a parser might have actions
to process versions one and two of a file format; the parser will then
parse either version one or version two depending on the ruleset selected.
The architecture makes it possible to apply commonly used programming
techniques such as object orientation, inheritance, composition, delegation
and roles when designing and implementing a parser, simplifying the process
of working within a team with shared responsibility or developing and
testing additional functionality.  This architecture is ideally suited to
parsing inputs where the input is not fully understood or does not follow a
fixed grammar: the architecture warns about unparsed inputs and other
errors, but continues on parsing as best it can, allowing the developer of
a new parser to decide which deficiencies are most important and require
attention first, rather than being forced to fix the first error which
arises.

The dataset gathered by the parser provides the foundation for the future
of this project: applying machine-learning algorithms to the data to
analyse and optimise the set of rules in use, followed by identifying
patterns in the data which could be used to write new filters to recognise
and reject spam rather than accepting it.  The parser provides the data in
a normalised form which is far easier to use as input to new or existing
implementations of the machine-learning algorithms than trying to adapt the
algorithms to extract data directly from the log files.  The current focus
is on clustering and decision trees to optimise the order in which rules
are applied; future efforts will involve using data gathered by the parser
to train and test new filters.  This task is very similar to optimising a
black box application based on its inputs and outputs, and this approach
could be applied to optimising the behaviour of any system given sufficient
logging to analyse.  An alternative approach to black box optimisation
which uses application profiling in conjunction with the application's
error messages to improve the error messages shown to users is described
in~\cite{black-box-error-reporting}; profiling data may be useful in
supplementing systems which fail to provide adequate log messages.
